---
title: "Data_Integration&Evaluation"
author: "Krithika Umesh"
date: "2023-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(Matrix)
library(harmony)
library(patchwork)
library(kBET)
library(cluster, quietly = TRUE)
library(liger)
library(SeuratWrappers)
library(lisi)
library(cowplot)
##Python env
Sys.setenv(RETICULATE_PYTHON = "C:\\Users\\krith\\anaconda3/python.exe")
library(reticulate)
sc <- import('scanpy', convert=FALSE)
scvi <- import("scvi", convert = FALSE)
scanorama <- import("scanorama")
library(sceasy)
library(anndata)
```

**Read individual data objects**

```{r, read individual data objects, echo=FALSE, cache=TRUE}
normal_1 <- readRDS("C:/Users/krith/Desktop/GSE162631/Objects/normal_1.rds")

tumor_1 <- readRDS("C:/Users/krith/Desktop/GSE162631/Objects/tumor_1.rds")

normal_2 <- readRDS("C:/Users/krith/Desktop/GSE162631/Objects/normal_2.rds")

tumor_2 <- readRDS("C:/Users/krith/Desktop/GSE162631/Objects/tumor_2.rds")

```

**Merge datasets and filter low quality cells**

```{r, merge datasets, echo=FALSE, cache=FALSE}
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"

VlnPlot(merge_data, features = c("nFeature_RNA", "nCount_RNA","percent.mt"), ncol=3)
plot1 <- FeatureScatter(merge_data, feature1 = 'nCount_RNA', feature2 = 'percent.mt')
plot2 <- FeatureScatter(merge_data, feature1 = 'nCount_RNA', feature2 = 'nFeature_RNA')
plot1+plot2

```

**Normalize and select variable features**

```{r, Normalize and feature selection,echo=FALSE, cache=TRUE}
batch_list <- SplitObject(merge_data, split.by = 'orig.ident')
# normalize and identify variable features for each dataset independently
batch_list <- lapply(X = batch_list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 5000)
})
# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = batch_list)

```

**Seurat Integration**

```{r, Integration, echo=FALSE, cache=TRUE}
anchors <- FindIntegrationAnchors(object.list = batch_list, anchor.features = features, k.filter= NA)
# this command creates an 'integrated' data assay
Seu_Int <- IntegrateData(anchorset = anchors)

##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

Seu_Int@meta.data$groups[Seu_Int@meta.data$orig.ident %in% Normal] <- "Normal"
Seu_Int@meta.data$groups[Seu_Int@meta.data$orig.ident %in% Tumor] <- "Tumor"

```

**PCA and elbow plot - Seurat**

```{r, Integrated Analysis,echo=FALSE, cache=TRUE}
DefaultAssay(Seu_Int) <- "integrated"
# Run the standard workflow for visualization and clustering
Seu_Int <- ScaleData(Seu_Int, verbose = FALSE)
Seu_Int <- RunPCA(Seu_Int, npcs = 50, verbose = FALSE)
ElbowPlot(Seu_Int, ndims=ncol(Embeddings(Seu_Int, "pca")))
print(Seu_Int[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(Seu_Int, dims = 1:2, reduction = "pca")

```

**UMAP**

```{r, Non-linear reduction,echo=FALSE, cache=TRUE}
Seu_Int <- RunUMAP(Seu_Int, reduction = "pca", dims = 1:50)
Seu_Int <- FindNeighbors(Seu_Int, reduction = "pca", dims = 1:50)
Seu_Int <- FindClusters(Seu_Int, resolution = c(0.01,0.05,0.1,0.5))

```

**Plots to view clusters**

```{r, Plots, echo=FALSE, cache=TRUE}
p1 <- DimPlot(Seu_Int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(Seu_Int, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2

p3 <- DimPlot(Seu_Int, reduction = "umap", group.by = "groups")
p3+p2

##UMAP Plots of 4 genes
plot_a <- UMAPPlot(Seu_Int, group.by="orig.ident")
plot_b <- UMAPPlot(Seu_Int, label=T)
plot_c <- FeaturePlot(Seu_Int, c("TMEM204", "VWF", "CLDN5", "CAVIN2"))
((plot_a+plot_b)|plot_c)+plot_layout(width=c(5,5))

```

**Seurat - Evaluation**

```{r, Seurat Evaluation, cache=TRUE, echo=FALSE}
##kBET
data <- GetAssayData(Seu_Int, slot = "scale.data")
data <- t(data)
data <- as.matrix(data)
batch <- Seu_Int$groups
batch.estimate <- kBET(data, batch, plot=TRUE)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = Seu_Int[['umap']]))
clusters <- Seu_Int$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
Seu_Int$sil <- sil[, 3]
print("The average silhouette score is:")
summary(Seu_Int$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = Seu_Int, grouping.var = "groups", reduction = 'pca', dims = 1:2, k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
Y <- Embeddings(Seu_Int[["pca"]])
Y <- as.matrix(Y)
metadata <- Seu_Int@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,6,7,8,9,10,11,12,13))
lisi_umap <- compute_lisi(Y, metadata, 'groups', perplexity=30)
print("The summary of lisi score is:")
summary(lisi_umap)

# Local structure preservation
DefaultAssay(object = Seu_Int) <- "RNA"
ls <- LocalStruct(object = Seu_Int, grouping.var = "groups", reduction = 'pca', orig.dims = 1:30, neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)
```

**LIGER integration**

```{r, Data Integration-LIGER, cache=TRUE, echo=FALSE}
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"

liger_Int <- merge_data %>% NormalizeData() %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 5000)
liger_Int <- ScaleData(liger_Int, split.by = 'orig.ident', do.center=FALSE)
liger_Int <- RunOptimizeALS(liger_Int, k=20, lambda = 5, split.by = 'orig.ident')
liger_Int <- RunQuantileAlignSNF(liger_Int, split.by = 'orig.ident')
liger_Int <- RunUMAP(liger_Int, dims=1:ncol(liger_Int[["iNMF"]]), reduction="iNMF")
liger_Int <- FindNeighbors(liger_Int, reduction="iNMF", dims=1:ncol(Embeddings(liger_Int, "iNMF"))) %>% FindClusters(resolution=0.4)

##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

liger_Int@meta.data$groups[liger_Int@meta.data$orig.ident %in% Normal] <- "Normal"
liger_Int@meta.data$groups[liger_Int@meta.data$orig.ident %in% Tumor] <- "Tumor"

```

**LIGER - Plots to view clusters**

```{r, Plots-Liger, cache=TRUE, echo=FALSE}
p1 <- DimPlot(liger_Int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(liger_Int, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2

p3 <- DimPlot(liger_Int, reduction = "umap", group.by = "groups")
p3+p2
##UMAP Plots of 4 genes
plot_a <- UMAPPlot(liger_Int, group.by="orig.ident")
plot_b <- UMAPPlot(liger_Int, label=T)
plot_c <- FeaturePlot(liger_Int, c("TMEM204", "VWF", "CLDN5", "CAVIN2"))
((plot_a+plot_b)|plot_c)+plot_layout(width=c(5,5))

```

**LIGER - Evaluation**

```{r, Evaluate-Liger, cache=TRUE, echo=FALSE}
##kBET
data <- GetAssayData(liger_Int, slot = "scale.data")
data <- t(data)
data <- as.matrix(data)
batch <- liger_Int$groups
batch.estimate <- kBET(data, batch, plot=TRUE, do.pca=TRUE, dim.pca=10)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = liger_Int[['iNMF']]))
clusters <- liger_Int$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
liger_Int$sil <- sil[, 3]
print("The average silhouette score is:")
summary(liger_Int$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = liger_Int, grouping.var = "orig.ident", reduction = 'iNMF', k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
Y <- Embeddings(liger_Int[["iNMF"]])
Y <- as.matrix(Y)
metadata <- liger_Int@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,5,6,7,9))
lisi <- compute_lisi(Y, metadata, 'groups', perplexity=30)
summary(lisi)

# Local structure preservation
DefaultAssay(object = liger_Int) <- "RNA"
ls <- LocalStruct(object = liger_Int, grouping.var = "groups", reduction = 'iNMF', neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)
```

**Harmony - integration**

```{r, Harmony integration, cache=TRUE, echo=FALSE}
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"

# Perform log-normalization and feature selection, scaling and PCA
harmony_Int <- merge_data %>% NormalizeData() %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 5000) %>% ScaleData() %>% RunPCA(npcs=50)
harmony_Int <- RunHarmony(harmony_Int, group.by.vars = "orig.ident", dims.use = 1:50, max.iter.harmony = 50)
harmony_Int <- RunUMAP(harmony_Int, reduction='harmony', dims=1:50)
harmony_Int <- FindNeighbors(harmony_Int, reduction='harmony', dims=1:50) %>% FindClusters(resolution=0.2)

##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

harmony_Int@meta.data$groups[harmony_Int@meta.data$orig.ident %in% Normal] <- "Normal"
harmony_Int@meta.data$groups[harmony_Int@meta.data$orig.ident %in% Tumor] <- "Tumor"

```

**Plots - Harmony**

```{r, Plots-Harmony, cache=TRUE, echo=FALSE}
##elbow plot and top genes in PC1 and PC2
ElbowPlot(harmony_Int, ndims=ncol(Embeddings(harmony_Int, "pca")))
VizDimLoadings(harmony_Int, dims = 1:2, reduction = "pca")

p1 <- DimPlot(harmony_Int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(harmony_Int, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2

p3 <- DimPlot(harmony_Int, reduction = "umap", group.by = "groups")
p3+p1
p1+p2+p3+ plot_layout(width=c(10,5))

##UMAP Plots of 4 genes
plot_a <- UMAPPlot(harmony_Int, group.by="orig.ident")
plot_b <- UMAPPlot(harmony_Int, label=T)
plot_c <- FeaturePlot(harmony_Int, c("TMEM204", "VWF", "CLDN5", "CAVIN2"))
((plot_a+plot_b)|plot_c)+plot_layout(width=c(5,5))

```

**Evaluation - Harmony**

```{r, Harmony-Evaluation, echo=FALSE, cache=TRUE}
##kBET
data <- GetAssayData(harmony_Int, slot = "scale.data")
data <- t(data)
data <- as.matrix(data)
batch <- harmony_Int$groups
batch.estimate <- kBET(data, batch, plot=TRUE, do.pca=TRUE, dim.pca=10)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = harmony_Int[['umap']]))
clusters <- harmony_Int$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
harmony_Int$sil <- sil[, 3]
print("The average silhouette score is:")
summary(harmony_Int$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = harmony_Int, grouping.var = "groups", reduction = 'pca', dims = 1:2, k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
Y <- Embeddings(harmony_Int[["umap"]])
Y <- as.matrix(Y)
metadata <- harmony_Int@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,5,6,8))
lisi_umap <- compute_lisi(Y, metadata, 'groups', perplexity=30)
print("The lisi summary is:")
summary(lisi_umap)

# Local structure preservation
DefaultAssay(object = harmony_Int) <- "RNA"
ls <- LocalStruct(object = harmony_Int, grouping.var = "groups", reduction = 'pca', reduced.dims = 1:30, orig.dims = 1:30, neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)

```

**RPCA - integration**

```{r, rpca integrate, cache=TRUE, echo=FALSE}
##load the data
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"

#Split object
batch_list <- SplitObject(merge_data, split.by = 'orig.ident')
# normalize and identify variable features for each dataset independently
batch_list <- lapply(X = batch_list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 5000)
})
# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = batch_list)
batch_list <- lapply(X = batch_list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})

```

```{r, perform rpca integration, cache=TRUE, echo=FALSE}

anchors <- FindIntegrationAnchors(object.list = batch_list, anchor.features = features, reduction = "rpca")

# this command creates an 'integrated' data assay
rpca_Int <- IntegrateData(anchorset = anchors, k.weight = 49)
```

```{r, analysis rpca, cache=TRUE, echo=FALSE}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(rpca_Int) <- "integrated"

# Run the standard workflow for visualization and clustering
rpca_Int <- ScaleData(rpca_Int, verbose = FALSE)
rpca_Int <- RunPCA(rpca_Int, npcs = 30, verbose = FALSE)
rpca_Int <- RunUMAP(rpca_Int, reduction = "pca", dims = 1:30)
rpca_Int <- FindNeighbors(rpca_Int, reduction = "pca", dims = 1:30)
rpca_Int <- FindClusters(rpca_Int, resolution = 0.5)

##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

rpca_Int@meta.data$groups[rpca_Int@meta.data$orig.ident %in% Normal] <- "Normal"
rpca_Int@meta.data$groups[rpca_Int@meta.data$orig.ident %in% Tumor] <- "Tumor"

```

**RPCA - Plots**

```{r, Plots-RPCA, cache=TRUE, echo=FALSE}
p1 <- DimPlot(rpca_Int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(rpca_Int, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2

p3 <- DimPlot(rpca_Int, reduction = "umap", group.by = "groups")
p3+p2
##UMAP Plots of 4 genes
plot_a <- UMAPPlot(rpca_Int, group.by="orig.ident")
plot_b <- UMAPPlot(rpca_Int, label=T)
plot_c <- FeaturePlot(rpca_Int, c("TMEM204", "VWF", "CLDN5", "CAVIN2"))
((plot_a+plot_b)|plot_c)+plot_layout(width=c(5,5))

```

**RPCA - Evaluation**

```{r, evaluate rpca, cache=TRUE, echo=FALSE}
##kBET
data <- GetAssayData(rpca_Int, slot = "scale.data")
data <- t(data)
data <- as.matrix(data)
batch <- rpca_Int$groups
batch.estimate <- kBET(data, batch, plot=TRUE, do.pca=TRUE, dim.pca=10)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = rpca_Int[['umap']]))
clusters <- rpca_Int$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
rpca_Int$sil <- sil[, 3]
print("The average silhouette score is:")
summary(rpca_Int$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = rpca_Int, grouping.var = "groups", reduction = 'pca', dims = 1:2, k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
Y <- Embeddings(rpca_Int[["umap"]])
Y <- as.matrix(Y)
metadata <- rpca_Int@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,5,6,8))
lisi_umap <- compute_lisi(Y, metadata, 'groups', perplexity=30)
print("The lisi summary is:")
summary(lisi_umap)

# Local structure preservation
DefaultAssay(object = rpca_Int) <- "RNA"
ls <- LocalStruct(object = rpca_Int, grouping.var = "groups", reduction = 'pca', reduced.dims = 1:30, orig.dims = 1:30, neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)
```

**scVI - integration**

```{r, scvi integration, echo=FALSE, cache=TRUE}
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"

##Variable gene selection
merge_data <- NormalizeData(merge_data, normalization.method = "LogNormalize", scale.factor = 10000)

merge_data <- subset(merge_data, subset = nFeature_RNA > 200 & nFeature_RNA < 8000)
merge_data <- FindVariableFeatures(merge_data, selection.method = "vst", nfeatures = 5000)
top5000 <- head(VariableFeatures(merge_data), 5000)
merge_data <- merge_data[top5000]

```

```{r, conversion to anndata, cache=TRUE}
adata <- convertFormat(merge_data, from="seurat", to="anndata", main_layer="counts", drop_single_values=FALSE)
print(adata)

```

```{r, train anndata model}
# run setup_anndata, use column stim for batch
scvi$model$SCVI$setup_anndata(adata, batch_key = 'orig.ident')

# create the model
model = scvi$model$SCVI(adata)

# train the model
model$train()

```

```{r, latent rep, cache=TRUE, echo=FALSE}
# get the latent represenation
latent = model$get_latent_representation()

# put it back in our original Seurat object
latent <- as.matrix(latent)
rownames(latent) = colnames(merge_data)
merge_data[["scvi"]] <- CreateDimReducObject(embeddings = latent, key = "scvi_", assay = DefaultAssay(merge_data))

##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

merge_data@meta.data$groups[merge_data@meta.data$orig.ident %in% Normal] <- "Normal"
merge_data@meta.data$groups[merge_data@meta.data$orig.ident %in% Tumor] <- "Tumor"

```

**Plots - scVI**

```{r, scvi Plots}
#library(cowplot)
# for jupyter notebook
#options(repr.plot.width=10, repr.plot.height=8)

merge_data <- RunUMAP(merge_data, dims = 1:10, reduction = "scvi", n.components = 2)
p1 <- DimPlot(merge_data, reduction = "umap", group.by = "orig.ident", pt.size=2)
plot_grid(p1)

p2 <- DimPlot(merge_data, reduction = "umap", group.by = "groups", pt.size=2)
plot_grid(p2)

```

**Evaluation - scVI**

```{r, scvi evaluation, cache=TRUE, echo=FALSE}
##kBET
data <- GetAssayData(merge_data, slot = "scale.data")
data <- t(data)
data <- as.matrix(data)
batch <- merge_data$groups
batch.estimate <- kBET(data, batch, plot=TRUE)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = merge_data[['scvi']]))
clusters <- merge_data$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
merge_data$sil <- sil[, 3]
print("The average silhouette score is:")
summary(merge_data$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = merge_data, grouping.var = "groups", reduction = 'scvi', dims = 1:2, k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
X <- Embeddings(merge_data[["scvi"]])
X <- as.matrix(X)
metadata <- merge_data@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,6))
lisi_scvi <- compute_lisi(X, metadata, 'groups', perplexity=30)
summary(lisi_scvi)

# Local structure preservation
DefaultAssay(object = merge_data) <- "RNA"
ls <- LocalStruct(object = merge_data, grouping.var = "groups", reduction = 'scvi', neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)
```

**Scanorama - integration**

```{r, merge and preprocess data for scanorama}
merge_data <- merge(x=normal_1, y=c(normal_2, tumor_1, tumor_2), add.cell.ids = c("Normal_1","Normal_2","Tumor_1","Tumor_2"), project="Combined")

Idents(merge_data) <- "orig.ident"
##subset to avoid running out of memory
#merge_data_sub <- subset(merge_data, downsample=1000)
batch_list <- SplitObject(merge_data, split.by = 'orig.ident')
# normalize and identify variable features for each dataset independently
batch_list <- lapply(X = batch_list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 5000)
})

```

```{r, create lists of matrices and genes, cache=TRUE, echo=FALSE}

assaylist <- list()
genelist <- list()
for(i in 1:length(batch_list))
{
   assaylist[[i]] <- t(as.matrix(GetAssayData(batch_list[[i]], "data")))
   genelist[[i]] <- rownames(batch_list[[i]])
}

```

```{r, scanorama integration, cache=TRUE, echo=FALSE}

integrated.data <- scanorama$integrate(assaylist, genelist)
corrected.data <- scanorama$correct(assaylist, genelist, return_dense=TRUE)
integrated.corrected.data <- scanorama$correct(assaylist, genelist, return_dimred=TRUE, return_dense=TRUE)
```

```{r, processing scanorama integrated object}
#extract from the objects we just created the corrected counts (element number 2), and the dimensional reduction embeddings (element number 1), and collate each of them together.

intdata <- lapply(integrated.corrected.data[[2]], t)
panorama <- do.call(cbind, intdata)

#use the common genes (element number 3) to give rownames to the integrated, batch-corrected expression matrix.
rownames(panorama) <- as.character(integrated.corrected.data[[3]])
colnames(panorama) <- unlist(sapply(assaylist, rownames))

intdimred <- do.call(rbind, integrated.corrected.data[[1]])
colnames(intdimred) <- paste0("PC_", 1:100)

   #We also add standard deviations in order to draw Elbow Plots in Seurat

stdevs <- apply(intdimred, MARGIN = 2, FUN = sd)

```

```{r, create seurat object from scanorama integrated}

pan.int <- CreateSeuratObject(counts = panorama, assay = "pano",  project = "Integrated_scanorama")

  #Adding metadata from all previous objects 
pan.int@meta.data <- do.call(rbind, lapply(batch_list, function(x) x@meta.data))
  
    # VERY IMPORTANT: make sure that the rownames of your metadata slot 
    # are the same as the colnames of your integrated expression matrix 

rownames(pan.int@meta.data) <- colnames(pan.int)
rownames(intdimred) <- colnames(pan.int)

pan.int[["pca"]] <- CreateDimReducObject(embeddings = intdimred, stdev = stdevs, key = "PC_", assay = "pano")
```

```{r, UMAP pano, cache=TRUE, echo=FALSE}
#Create elbow plot
ElbowPlot(pan.int, ndims=ncol(Embeddings(pan.int, "pca")))
print(pan.int[["pca"]], dims = 1:5, nfeatures = 5)
#VizDimLoadings(pan.int, reduction = "pca")

pan.int <- RunUMAP(pan.int, reduction = "pca", dims = 1:50)
pan.int <- FindNeighbors(pan.int, reduction = "pca", dims = 1:50)
pan.int <- FindClusters(pan.int, resolution = 0.6)
```

**Plots - Scanorama**

```{r, plots scanorama, echo=FALSE, cache=TRUE}
##Create 2 groups: Normal and Tumor
Normal <- c("Normal_1", "Normal_2")
Tumor <- c("Tumor_1", "Tumor_2")

pan.int@meta.data$groups[pan.int@meta.data$orig.ident %in% Normal] <- "Normal"
pan.int@meta.data$groups[pan.int@meta.data$orig.ident %in% Tumor] <- "Tumor"


##Plots
p1 <- DimPlot(pan.int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(pan.int, reduction = "umap", label = TRUE, repel = TRUE)
p1+p2

p3 <- DimPlot(pan.int, reduction = "umap", group.by = "groups")

p3+p1

##UMAP Plots of 4 genes
plot_a <- UMAPPlot(pan.int, group.by="orig.ident")
plot_b <- UMAPPlot(pan.int, label=T)
plot_c <- FeaturePlot(pan.int, c("TMEM204", "VWF", "CLDN5", "CAVIN2"))
((plot_a+plot_b)|plot_c)+plot_layout(width=c(5,5))
```

**Evaluation - Scanorama**

```{r, Evaluation, scanorama, cache=TRUE, echo=FALSE}

##kBET
data <- GetAssayData(pan.int, slot = "data")
data <- t(data)
data <- as.matrix(data)
batch <- pan.int$groups
batch.estimate <- kBET(data, batch, plot=TRUE)

##Silhouette score
dist.matrix <- dist(x = Embeddings(object = pan.int[['umap']]))
clusters <- pan.int$groups
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
pan.int$sil <- sil[, 3]
print("The average silhouette score is:")
summary(pan.int$sil)

# mixing metric
max.k <- 300
mm <- MixingMetric(object = pan.int, grouping.var = "groups", reduction = 'pca', dims = 1:2, k=5, max.k = max.k)
print("The average mixing metric score is:")
summary(mm)

#Compute LISI score
Y <- Embeddings(pan.int[["umap"]])
Y <- as.matrix(Y)
metadata <- pan.int@meta.data
metadata <- subset(metadata, select= -c(1,2,3,4,5,6))
lisi <- compute_lisi(Y, metadata, 'groups', perplexity=30)
print("The summary of lisi score is:")
summary(lisi)

# Local structure preservation
#DefaultAssay(object = pan.int) <- "RNA"
ls <- LocalStruct(object = pan.int, grouping.var = "groups", reduction = 'pca', orig.dims = 1:30, neighbors = 100)
ls <- unname(obj = unlist(x = ls))
print("The average score for local structure preservation is:")
summary(ls)
```

## R Markdown
